---
title: "PROJECT"
author: "Adeola Mary Olaniyan"
date: "2/29/2020"
output: html_document
---

```{r}
myeloma.data <- read.csv("C:/Users/adeol/Downloads/myeloma-data.csv", header=FALSE)

names(myeloma.data) <- c("patient-number", "stime", "status", "age", "sex", "bun", "ca", "hb", "p-cells", "protein")

library(survival)
library(tidyverse)
head(myeloma.data)
dim(myeloma.data)
fit.1 <- survfit(Surv(stime,status)~sex,conf.type="log-log", data = myeloma.data)
summary(fit.1)
```

3i. comparing the survival experience for the sex, where, male is 1 and female is 2.
the logrank test comparing the survival experience of the two groups

Let   H0:h1(t) =h2(t)
Let   HA:h1(t)≠h2(t)
@ ∝=0.05
Test statistic
```{r}
survdiff(Surv(stime,status)~sex, rho= 0, data = myeloma.data )
```

Since P-value =0.8 is greater than 0.05, we therefore fail to reject the null hypothesis and conclude that there is no significant in the sex.

 the Gehan/Breslow generalized Wilcoxon test comparing the two groups sex
Let   H0:h1(t) =h2(t)
Let   HA:h1(t)≠h2(t)
@ ∝=0.05
Test statistic

```{r}
survdiff(Surv(stime,status)~sex, rho= 1, data = myeloma.data)
```

Since P-value =1 is greater than 0.05, we therefore fail to reject the null hypothesis and conclude that there is no significant in sex.

#ii. comparing the survival experience for the protein, for the two groups.
# the logrank test comparing the survival experience of the two groups.
Let   H0:h1(t) =h2(t)
Let   HA:h1(t)≠h2(t)
@ ∝=0.05
Test statistic
```{r}
survdiff(Surv(stime,status)~protein, rho= 0, data = myeloma.data )
```


Since P-value =0.2 is less than 0.05, we therefore reject the null hypothesis and conclude that there is significant in protein.

the Gehan/Breslow generalized Wilcoxon test comparing the two groups
Let   H0:h1(t) =h2(t)
Let   HA:h1(t)≠h2(t)
@ ∝=0.05
Test statistic


```{r}
survdiff(Surv(stime,status)~protein, rho= 1, data = myeloma.data)
```
Since P-value =0.1 is less than 0.05, we therefore reject the null hypothesis and conclude that there is significant in protein.

4.The maximum likelihood estimate of the probability of discontinuation by 52 weeks.
```{r}
library(readxl)
IUD_data1 <- read_excel("Book4.xlsx")
names(IUD_data1)<- c ("time","status")
fit.IUD_data1 <- survreg(Surv(time,status)~ 1, dist = "weibull",data = IUD_data1)
summary(fit.IUD_data1)
```


5 The ovarian cancer data set

```{r}
ovariancan <-  read.csv("C:/Users/adeol/Downloads/homework2.txt", sep="")

names(ovariancan)
age <-ovariancan[,1] 
grade <-ovariancan[,2] 
stage<-ovariancan[,3]
resi<-ovariancan[,4]
treatment<-ovariancan[,5]
time<-ovariancan[,6]
delta<-ovariancan[,7]
```

5a Non-parametric estimate of the survival function for each grade and  the plots to check graphically whether the exponential or the Weibull distributions seem reasonable.
```{r}
library(survminer)
library(survival)
allgrades<-survfit(Surv(time,delta)~grade,conf.type="log-log",data=ovariancan)
summary(allgrades)

ggsurvplot(allgrades)
ggsurvplot(allgrades, fun = "cumhaz")
```

Grade 1 has the best prognosis because the survival probability versus time graph shows that the survival time is higher, while Grade 3 has the worst prognosis because the same graph indicates that the survival time is lower.

The exponential distribution seems reasonable because the cummulative hazard graph is linear to time.

5b. An exponential model separately within each grade.


```{r}
library(survival)

grade1 <- ovariancan[ovariancan$grade == 1,]
grade2 <- ovariancan[ovariancan$grade == 2,]
grade3 <- ovariancan[ovariancan$grade == 3,]



grade1exp <- survreg(Surv(time,delta)~1, dist = "exponential", data=grade1)
summary(grade1exp)

grade2exp <- survreg(Surv(time,delta)~1, dist = "exponential", data=grade2)
summary(grade2exp)

grade3exp <- survreg(Surv(time,delta)~1, dist = "exponential", data=grade3)
summary(grade3exp)
```



5c. the exponential model, test at the α = 0.05 level of significance

Let   H0:h1(t) =h2(t)=h3(t)
Let   HA: at least one of the grades has different distribution of time to reccurrence
@ ∝=0.05
Test statistic


```{r}
gradetest <- survreg(Surv(time,delta)~grade, dist = "exponential", data=ovariancan)
summary(gradetest)
```


 p-values 
Since P-value =0.43 is greater than 0.05, we therefore fail to reject the null hypothesis and conclude that there is no significant in distribution of time to  recurrence.
 the likelihood ratio test
the likelihood ratio is -410.3
the Wald test

d.using a Weibull model
the estimated hazard functions and corresponding 95% confidence intervals.
```{r}
library(SurvRegCensCov)

gradeweib <- survreg(Surv(time,delta)~grade, dist = "weibull", data=ovariancan)
summary(gradeweib)

ConvertWeibull(gradeweib,conf.level = 0.95)

grade1weib <- survreg(Surv(time,delta)~1, dist = "weibull", data=grade1)
ConvertWeibull(gradeweib,conf.level = 0.95)
summary(grade1weib)

grade2weib <- survreg(Surv(time,delta)~1, dist = "weibull", data=grade2)
ConvertWeibull(gradeweib,conf.level = 0.95)
summary(grade2weib)

grade3weib <- survreg(Surv(time,delta)~1, dist = "weibull", data=grade3)
ConvertWeibull(gradeweib,conf.level = 0.95)
summary(grade3weib)

summary(gradeweib$coefficients)

```
the estimated mean survival time is 4.46887 and the median of the survial time is 4.46887
